Review Paper Draft

### 2020-06-25 main.tex
Figures: tweaks/notes to self

The mechanism of environmental confounding is summarized in the directed acyclic graph displayed in figure \ref{fig:ps_env}. The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction. Otherwise stated, we do not assume that environmental factors influence population-specific allele frequencies, or vice versa, only that the two may be related. Additionally, we assume that the environmental exposure has no effect on on observed SNPs.  \pb{I find these previous two sentences confusing/contradictory: We assume that environment and genetics are related, but also that environment has no effect on SNPs.}

@@

Statistical methods to control for the effects of confounding due to population stratification have proliferated as computational advances have allowed for the analysis of genetic data from large cohorts. It should be noted that by attempting to reduce the effects of population structure, these methods reduce the effects of environmental confounding by blocking the casual pathway between population structure and observed SNPs \pb{Shouldn't this be: between environment and SNPs?}.

@@

Pseudophenotypes for these simulations were again generated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$. Figure \ref{fig:big_vs_small} summarizes the results of these simulations with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures. The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy. In the presence of a small number of large subpopulations, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding box straddling the 0 line. However, the majority of the box falls below 0, indicating the superior performance of PC-Lasso in many of the simulations. In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy. \cite{hoffman2013correcting} showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models. As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations.

\anna{admixture simulated data sets do worse in the large subpops setting because their similarity matrix does not indicate as clearly who was exposed to one of e.g. 4 environmental exposures. the small subpopulations scenario more closely mimics gradations of environmental effects - what is usually simulated with a random draw from the MVN distn... we're simulating environmental heterogeneity in a way that allows us to examine the bias produced and these methods still do as well or better than pca}

>>>>

The mechanism of environmental confounding is summarized in the directed acyclic graph displayed in figure \ref{fig:ps_env}. The undirected dashed line between population stratification and environmental exposure indicates that while we assume these elements are correlated, we do not assume any causal direction. Otherwise stated, we do not assume that environmental factors influence population-specific allele frequencies, or vice versa, only that the two may be related. Additionally, we assume that the environmental exposure has no \anna{direct} effect on on observed SNPs.  \pb{I find these previous two sentences confusing/contradictory: We assume that environment and genetics are related, but also that environment has no effect on SNPs.} \anna{we assume the environment is correlated with population structure, and thus genetic information may serve as a proxy for differential environmental exposures, which also affect phenotype}

@@

Statistical methods to control for the effects of confounding due to population stratification have proliferated as computational advances have allowed for the analysis of genetic data from large cohorts. It should be noted that by attempting to reduce the effects of population structure, these methods reduce the effects of environmental confounding by blocking the casual pathway between population structure and observed SNPs \pb{Shouldn't this be: between environment and SNPs?}. \anna{no...}

@@

Pseudophenotypes for these simulations were again generated using a 1:1 SNR, $\eta = 0.5$, and in the presence of moderate environmental confounding, $\xi = 0.5$. Figure \ref{fig:big_vs_small} summarizes the results of these simulations with the y-axis displaying the difference in MSE (PC-Lasso - LMM-Lasso) for each simulation and for the various subpopulation structures. The dashed horizontal line at 0 indicates no difference in the performance of PC-Lasso and LMM-Lasso, while differences greater than 0 correspond to smaller MSE for LMM-Lasso, and thus superior estimation accuracy. In the presence of a small number of large subpopulations, there is some overlap in the estimation accuracy of PC-Lasso and LMM-Lasso, as evidenced by the corresponding box straddling the 0 line. However, the majority of the box falls below 0, indicating the superior performance of PC-Lasso in many of the simulations. In the presence of many small subpopulations, however, LMM-Lasso is clearly superior to PC-Lasso in estimation accuracy. \cite{hoffman2013correcting} showed that principal component adjustment methods represent a low dimensional approximation to linear mixed models. As such, it is reasonable that PC-Lasso, which corrects for population structure using a low dimensional set of surrogate variables, would perform well in the small-number-of-large-subpopulations setting, but fall short in the presence of many small subpopulations. \\

\anna{
admixture simulated data sets do worse in the large subpops setting because their similarity matrix does not indicate as clearly who was exposed to one of e.g. 4 environmental exposures; it is a better proxy for environmental effects.\\
}

\anna{
the small subpopulations scenario more closely mimics a gradient of environmental effects - what is usually simulated with a random draw from the MVN distn... \\
}

\anna{
we're simulating environmental heterogeneity in a way that allows us to examine the bias produced and these methods still do as well or better than pca}\\
